---
title: "BDA - Project"
author: "Anonymous"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 1
urlcolor: blue
---

```{r setup, include=FALSE}
# This chunk sets echo = TRUE as default, that is print all code.
# knitr::opts_chunk$set can be used to set other notebook generation options, too.
# include=FALSE inside curly brackets makes this block not be included in the pdf.
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# To install aaltobda, see the General information in the assignment.
library(aaltobda)
library(pracma)
library(posterior)
library(cmdstanr)
library(rstan)
library(ggplot2)
library(loo)
library(dplyr)
cmdstanr::check_cmdstan_toolchain(fix = TRUE)
data("factory")

# SEED = 48927

register_knitr_engine(override = FALSE)
```

# Introduction

## Motivation

## Problem

# Data

## Some EDA ----

```{r}
## finish position
ggplot(f1_dat_finished, aes(x = factor(position))) +
  geom_bar() +
  labs(
    title = "Distribution of finish positions",
    subtitle = "F1 hybrid era (2014-2020)",
    x = "Finish position",
    y = "Count"
  )
```

```{r}
f1_dat_finished %>%
  filter(driver %in% c("hamilton", "raikkonen", "giovinazzi")) %>%
  ggplot(aes(x = factor(position), fill = driver)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  labs(
    x = "Finish position",
    y = "Count",
    title = "Different drivers' finish positions",
    subtitle = "Conditional on finishing the race",
    fill = ""
  ) +
  theme(legend.position = "top") +
  facet_wrap(~year)
```

```{r}
f1_dat_finished %>%
  filter(driver %in% c("hamilton", "raikkonen", "giovinazzi"), year != 2014) %>%
  ggplot(aes(x = prop_trans, fill = driver)) +
  geom_density(alpha = 0.5, bw = 0.1) +
  labs(
    x = "Smoothed proportion of outperformed drivers",
    y = "Density",
    title = "Different drivers' results",
    subtitle = "Proportion of finished drivers outperformed",
    fill = ""
  ) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 45, vjust = 0.85)) +
  facet_wrap(~year)
```

# Describtion of the models

## Priors

# Stan models

# Rhat convergence diagnostics and interpretation

#  HMC specific convergence diagnostics

# Effective sample size diagnostic (n_eff)

# Posterior predictive checking and interpretation 

# Model comparison and interpretation of the results

# Predictive performance assessment

# Alternative priors testing

# Problems and potential improvements

# Conclusion

# Self-reflection about what the group learned 