---
title: "BDA - Project"
author: "Anonymous"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 1
urlcolor: blue
---

```{r setup, include=FALSE}
# This chunk sets echo = TRUE as default, that is print all code.
# knitr::opts_chunk$set can be used to set other notebook generation options, too.
# include=FALSE inside curly brackets makes this block not be included in the pdf.
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# To install aaltobda, see the General information in the assignment.
library(aaltobda)
library(pracma)
library(posterior)
library(cmdstanr)
library(rstan)
library(ggplot2)
library(loo)
library(dplyr)
cmdstanr::check_cmdstan_toolchain(fix = TRUE)
data("factory")

# SEED = 48927

register_knitr_engine(override = FALSE)
```

# Introduction

## Motivation

## Problem

# Data

## Some EDA ----

```{r}
data = read.csv("./data/race_results_view.csv")
```

```{r}
# Data processing 
## Restricting my analysis to the period 2012-2021
data <- data %>% filter(
  position > 0,
  year > 2011
)
## convert to factors
data <- data %>% mutate(
  rider_name  = as.factor(rider_name),
  team_name  = as.factor(team_name)
)

# New variables
data <- data %>% group_by(year, sequence) %>% mutate(  
  position_prop = (n() - position) / (n() - 1),        
  prop_trans = (position_prop * (n() - 1) + 0.5) / n() 
  )
```
The variable position_prop represents how many rider_names you beat in a race, the variables position_trans is a transformation (see https://stats.stackexchange.com/a/134297/116878)

Preliminary plots


```{r}
## finish position
ggplot(data, aes(x = factor(position))) +
  geom_bar(fill = "darkmagenta") +
  labs(
    title = "Distribution of finish positions",
    subtitle = "Era (2012-2021)",
    x = "Finish position",
    y = "Count"
  )
```
Dire che non ce ne sono sempre 22 perchè non tutti arrivano in fondo

```{r}
data %>%
  filter(rider_name %in% c("Rossi, Valentino", "Quartararo, Fabio", "Marquez, Marc", "Lorenzo, Jorge")) %>%
  ggplot(aes(x = factor(position), fill = rider_name)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_x_discrete(limits=rev,breaks=seq(1, 23, 3)) +
  labs(
    x = "Finish position",
    y = "Count",
    title = "Different rider's finish positions",
    subtitle = "Conditional on finishing the race",
    fill = ""
  ) +
  theme(legend.position = "top") +
  facet_wrap(~year)
```
Dire che non tutti i piloti sono sempre presenti (alcuni si ritirano, altri arrivano dopo)

```{r}
data %>%
  filter(rider_name %in% c("Rossi, Valentino", "Crutchlow, Cal", "Marquez, Marc")) %>%
  ggplot(aes(x = prop_trans, fill = rider_name)) +
  geom_density(alpha = 0.5, bw = 0.1) +
  labs(
    x = "Smoothed proportion of outperformed rider_names",
    y = "Density",
    title = "Different rider_names' results",
    subtitle = "Proportion of finished drivers outperformed",
    fill = ""
  ) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 45, vjust = 0.85)) +
  facet_wrap(~year)
```

# Describtion of the models

## Priors

# Stan models

# Rhat convergence diagnostics and interpretation

#  HMC specific convergence diagnostics

# Effective sample size diagnostic (n_eff)

# Posterior predictive checking and interpretation 

# Model comparison and interpretation of the results

# Predictive performance assessment

# Alternative priors testing

# Problems and potential improvements

# Conclusion

# Self-reflection about what the group learned 